\documentclass[11pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathpartir}
\usepackage{verbatim}
\usepackage{tabularx}
\setlength\parindent{0pt}
\newcommand\DOT{\textbf{dot}}
\newcommand\COLON{\textbf{colon}}
\newcommand\STAR{\textbf{star}}
\newcommand\IDENT{\textbf{ident}}
\newcommand\HAS{\textbf{has}}
\newcommand\IMPORT{\textbf{import}}
\newcommand\END{\textbf{end}}
\newcommand\nonterm[1]{\langle\textnormal{#1}\rangle}
\newcommand\expr[1]{|\,#1\;|}
\begin{document}
\title{The Module Language}
\author{Zachary Kieda (zkieda@andrew.cmu.edu)}
\date{\today}
\maketitle
\begin{abstract}
\setlength\parindent{0pt}\noindent 

At some point in time any project gets large enough that the developer[s] decide to design their own language that will aide in accomplishing their various routine or evil goals.\\

Here, we present a language used for own sinister goals -- creating a easier way to construct Java modules, especially with respect to our CrowdSimulation project. In our project, a module belongs in a hierarchical tree rooted at a player. Modules present an easy way to make additions to the player's functionality, as well as provide an easy way to construct a tree of dependent modules for testing. Ultimately, we will use the Module Language (ModLang) to construct and perform tests, as well as run the code for our crowd simulation project. We hope this language will provide a \emph{natural} mechanism for creating a module system in any project.\\

At the present moment, we have the lexer, parser, and dynamic semantics working for ModLang 1.0. We will be working towards ModLang 2.0 to provide additional functionality (like importing modules from other files, and a unification of the type system). \\

\textbf{Keywords.} Programming Languages, Modules
\end{abstract}
\pagebreak
\begin{section}{Module System (Dynamic Semantics)}
\begin{subsection}{Module Specification}
We construct modules in a tree-like system, so modules have two parameterized types, $\textsf{Node}$ and $\textsf{Leaf}$. We use this specification to define the dynamic semantics of the language.\\

The signature for the $\textsf{Node}$ module is specified below. \\

\qquad $\begin{array}{l}
\mbox{signature for }\textsf{T Node} = \\
\qquad  \begin{array}{lcl}
\textsf{add} & : & \textsf{T Node} \rightarrow (\textsf{T Node})\; \textsf{Module} \rightarrow T\\
\textsf{init} & : & \textsf{T Node} \rightarrow ()\\
\textsf{new} & : & () \rightarrow \textsf{T Node}
\end{array}
\end{array}$\\

We define $\textsf{T Module}$ such that $\textsf{T Node }\verb|instanceof|\textsf{ T Module}$ and \\
$\textsf{T Leaf }\verb|instanceof|\textsf{ T Module}$. This means that the $\textsf{add}$ function accepts
both $\textsf{Node}$ types and $\textsf{Leaf}$ types. For object-oriented programming languages, we assume that $\textsf{add}$ and $\textsf{init}$ are methods and the current object is the first argument for these functions.\\

The signature for the $\textsf{Leaf}$ module is specified below. \\

\qquad $\begin{array}{l}
\mbox{signature for }\textsf{T Leaf} = \\
\qquad  \begin{array}{lcl}
\textsf{init} & : & \textsf{T Leaf} \rightarrow ()\\
\textsf{new} & : & () \rightarrow \textsf{T Leaf}
\end{array}
\end{array}$\\

\textbf{Signature Contracts}.\,\\
We assert several contracts associated with these signatures that ensure that the modules work correctly. \\

For a module of type \textsf{T Node}, we list the method type, the contract type (requires/ensures), and the contract itself\\

\begin{tabularx}{\linewidth}{l l X}
$m.\textsf{init}()$ & requires & $\forall m'$ where $m'$ is ancestor of $m$, $m'.\textsf{init}()$ has been called. \\
& ensures & $\forall m'$ where $m'$ is a descendent of $m$, $m'.\textsf{init}()$ has completed.\\
$m.\textsf{add}(t)$ & requires & $t\;\texttt{instanceof}\;\textsf{T Module}$\\
& ensures & $t$ is a child of $m$
\end{tabularx}\\

A module of type $\textsf{T Leaf}$ has the same requires contract as $\textsf{T Node.init}$. However, the ensures contract only ensures that $\textsf{init}$ method has been called on this leaf.\\

\textbf{Operations.}\medskip\\
Additional operations can be provided in the system. For example, tree traversals that allow a module to communicate with its children/parent in either DFS or BFS ordering. 
\end{subsection}
\begin{subsection}{Java Implementation}
In our java implementation, the types
\begin{align*}
\textsf{T Leaf} &= \texttt{SubModule<T>}\\
\textsf{T Module} &= \texttt{MultiModule<T>}
\end{align*}
The $\textsf{new}$ method is equivalent to a constructor that takes no args. This means that all modules used in our language have a no-arg constructor.\\

We provide a utility annotation \verb|@AutoWired| that allows modules to easily access each other without 
setup. The annotation \verb|@AutoWired| can only be applied to fields of type \verb|Module|.\\

The code for this annotation is seen below
\begin{verbatim}
  @Retention(RUNTIME)
  @Target(FIELD)
  public @interface AutoWired {
    public String value() default "";
    public boolean noThrow() default true;
  }
\end{verbatim}

All fields of type $\tau : Module$ annotated with \verb|@AutoWired| will automatically find a class of type $\tau$ that is connected to this module tree and inject it into the annotated field. These fields are injected during the \emph{linking} phase.\\

If \verb|noThrow| is \verb|false|, we throw a Linking Exception if a module that extends $\tau$ could not be found at runtime. \\

The \verb|value| field allows the user to specify a specific location for the linker to look for an injected class. If $\verb|value| = \verb|""|$, then the linker will search all connected modules, and take the nearest child.\\

When the parameter \verb|value| is non-empty it either defines a relative path or an absolute path to the module we're looking for with respect to the module tree. Java classes in \verb|value| are separated by forward slashes. When \verb|value| begins with a forward slash, it is an absolute path and we search from the root module. Otherwise, we search relative to the current module. We can use \verb|"../"| to access the parent of this module. \\

Here's an example, 
\begin{verbatim}
  @AutoWired("../PerceptionModule/FieldOfVisionModule", 
    noThrow=false) 
  private FieldOfVisionModule fov;
\end{verbatim}
will search the parent module for a \verb|PerceptionModule|, then search that module for a \verb|FieldOfVision| module. If we can't find this module, then we throw a linking exception. Otherwise, the module we found is injected into field $\verb|fov|$.

\end{subsection}
\end{section}
\setcounter{section}{0}
\begin{center}\textbf{\LARGE Module Language 1.0} \end{center}

\begin{section}{Syntax}
\begin{subsection}{Tokens}
\textbf{Whitespace and Token Delimiting} 

\medskip Whitespace is either a space, a horizontal tab (\verb|\t|), vertical tab (\verb|\v|), linefeed (\verb|\n|), carriage return (\verb|\r|), or form feed (\verb|\f|) in ASCII encoding. Whitespace is ignored, except that it delimits tokens. Note that whitespace is not a \emph{requirement} to terminate a token. Whitespace can disambiguate two tokens when one of them is present as a prefix in another. The lecture should produce the longest valid token sequence possible.\\

The tokens and lexing is presented below in BNF, 
\[
\begin{array}{lcl}
\IDENT & ::= & \verb|[a-zA-Z$_][a-zA-Z0-9$_]*|\\
\COLON & ::= & \verb|:|\\
\DOT & ::= & \verb|.|\\
\STAR & ::= & \verb|*|\\
\IMPORT & ::= &\verb|import|\\
\HAS & ::= & \verb|has|\\
\END & ::= &\verb|end|
\end{array}
\]
Terminal referenced in the grammar are in \textbf{bold}. \IDENT\;is described using regular expressions.
\end{subsection}
\begin{subsection}{Grammar}
The context free grammar is presented below,
\[
\begin{array}{lcl}
\nonterm{program} & ::= & \nonterm{imports}\;\nonterm{module}\\
& | & \nonterm{module}\\
\nonterm{javaRef} & ::= & \IDENT\;\DOT\;\nonterm{javaRef}\\
& | & \IDENT\\
\nonterm{module} & ::= & \nonterm{javaRef}\\
& | & \nonterm{javaRef}\;\COLON\;\HAS\;\nonterm{moduleList}\;\END\\
& | & \nonterm{javaRef}\;\COLON\;\nonterm{module}\\
\nonterm{moduleList} & ::= & \nonterm{module}\;\nonterm{moduleList}\\
& | & \epsilon\\
\nonterm{moduleImport} & ::= & \IDENT\;\DOT\;\STAR\\
& | & \IDENT\\
& | & \IDENT\;\DOT\;\nonterm{moduleImport}\\
\nonterm{imports} & ::= & \IMPORT\;\nonterm{importList}\;\END\\
\nonterm{importList} & ::= & \nonterm{moduleImport}\;\nonterm{importList}\\
& | & \epsilon
\end{array}
\]

The parsed AST is returned as a $\nonterm{program}$. The $\nonterm{imports}$ section is used to import existing java modules from the project given by their absolute java path.
\end{subsection}
\begin{subsection}{Code Example}
We present a code example of ModLang that generates a player with the modules \verb|GameModule|, \verb|PerceptionModule|. The \verb|VieldOfVisionModule| is a child of the \verb|PerceptionModule|.
\begin{verbatim}
import
  edu.cmu.cs464.p3.ai.core.*
  edu.cmu.cs464.p3.ai.perception.*
end

Player : 
has
  GameModule
  PerceptionModule : FieldOfVisionModule
end
\end{verbatim}
In the first four lines we declare our imports. There is at most one import block, which must be declared before a module. The compiled script should return a single, fully initialized and linked module. It's important 
to note that this language is just markup and is not turing complete. This language is supposed to easily generate a hierarchy of modules to produce a module that will be used in a turing complete language. 
\end{subsection}
\end{section}
\begin{section}{Elaboration}
As the name is intended to suggest, \emph{elaboration} is the process of transforming the literal parse tree into to one that is simpler and more well behaved -- the abstract syntax tree. Much of this may be accomplished directly in the semantic actions that accompany the grammar rules.\\

We propose the following tree structure as the abstract syntax tree\\
\begin{align*}
\emph{prog} &::= 
%\emph{decl}\; |\;
\textsf{program}(\emph{imports}, \emph{mod})\\
%\emph{decl} &::= \emph{mod}\;|\;\textsf{declare}(\emph{import}, \emph{import})\\
\emph{imports} &::= \texttt{[]}\;|\;\emph{import} :: \emph{imports}\\
\emph{mod} & ::= \textsf{parent}(\emph{javaRef}, \emph{modList})\\
\emph{modList} & ::=\texttt{[]}\;|\;\emph{mod} :: \emph{modList}\\
\emph{import} & ::= \emph{package}\;\DOT\;\emph{javaRef} \;|\;\emph{package}\;\DOT\;\STAR
\end{align*}

We propose the following inference rules that describe how to elaborate the grammar into the abstract syntax tree.
\begin{mathpar}
\inferrule{\;}{\epsilon \leadsto \texttt{[]}} 

\inferrule{\nonterm{imports}\leadsto imp \\ \nonterm{module}\leadsto m}
  {\nonterm{imports}\;\nonterm{module}\leadsto \textsf{program}(imp, m)}
  
\inferrule{\nonterm{module} \leadsto m\\ \nonterm{moduleList}\leadsto M}
  {\nonterm{module}\;\nonterm{moduleList}\leadsto m \texttt{::} M} 
  
\inferrule{\nonterm{javaRef}\leadsto j\\ \nonterm{module}\leadsto m}
  {\nonterm{javaRef}\;\COLON\;\nonterm{module}\leadsto\textsf{parent}(j, [m])} 

\inferrule{\nonterm{javaRef}\leadsto j}
  {\nonterm{javaRef}\leadsto\textsf{parent}(j, [\;])} 


\inferrule{\nonterm{module}\leadsto m \\ \nonterm{moduleList}\leadsto M}
  {\nonterm{module} \;\COLON\;\HAS\;\nonterm{moduleList}\;\END \leadsto \textsf{parent}(m, M)} 

\inferrule{\nonterm{importList}\leadsto L}
  {\IMPORT\;\nonterm{importList}\;\END\leadsto L}

\inferrule{\nonterm{moduleImport}\leadsto l\\ \nonterm{importList}\leadsto L}
  {\nonterm{moduleImport}\;\nonterm{importList}\leadsto l \texttt{::} L}
\end{mathpar}
\end{section}
%To further express the import declaration semantics, we perform a different stage of elaboration using the function \emph{evalProg}, 
%\[\emph{evalProg}\;\textsf{program}(L, \emph{mod}) = 
%\begin{cases}
%\textsf{declare}(x, \emph{evalProg}\;\textsf{program}(X, \emph{mod})) & \mbox{if } L = x :: X\\
%\emph{mod} & \mbox{if } L = \texttt{[]}
%\end{cases}
%\]
%This generates a hierarchy of tree-like 
\begin{section}{Static Semantics}
In order to express the static semantics of this language, we use a unholy mashup of denotational semantics notation and classic static semantics notation.\\ 

Since we're dealing with classes in java, a type might have many super types because of inheritance. We write $\tau <: \tau'$ when the class $\tau$ is a subtype of, or extends the class $\tau'$. We define the type \textbf{Ty} to describe java classes, such that $\tau \in \textbf{Ty}$ if and only if $\tau <: \verb|Module|$.\\

We define the type $\textbf{Ns} = \emph{javaRef} \rightharpoonup \textbf{Ty}$ to define our namespace. The  partial function $\Gamma : \textbf{Ns}$ to express a namespace from a java reference to a java module class. Classically $\Gamma \vdash x : \tau\;\emph{valid}$ is written to express that expression $x$  has type $\tau$ under the context $\Gamma$. Here, we will write $(x, \tau) \in \Gamma$ as an equivalent expression.\\

The partial function $\verb|ClassLoader| : \emph{javaRef} \rightharpoonup \verb|Class<|?\verb|>|$ is used to access java classes by using their abstract path name. The \verb|ClassLoader| is a component of the JVM standard used to access classes reflectively. \\

We define the initial state of the program. The initial namespace, $\Gamma_i$, is 
\[(x, \tau) \in \Gamma_i \iff (x, \verb|Class<|\tau\verb|>|) \in \verb|ClassLoader|, \tau <: \verb|Module|\]

When parsing, we also keep track of a \emph{secondary} context, $\mathcal{C} : \textbf{Ty}_\bot$ where $\textbf{Ty}_\bot = \textbf{Ty}\cup\{\bot\}$. $\mathcal{C}$ keeps track of the current module type. This ensures that added submodules will also have a valid type with respect to their parent. Initially, $\mathcal{C} = \bot$ to indicate there is no current parent.\\

In order to formally define evaluation rules, we define the current state of the evaluation as $\textbf{St} = (\textbf{Ns}, \textbf{Ty}_\bot)$. For any statement $s$ in our AST, $\expr{e} : \textbf{St} \rightharpoonup \textbf{St}$ is a partial function. We say $(\sigma, \sigma') \in \expr{e}$ if and only if the expression $e$ will terminate with an ending state $\sigma'$ from the starting state $\sigma$.\\

Our rules are presented below,
\begin{mathpar}
\inferrule{\;}
  {(\sigma, \sigma) \in \texttt{[]}}

\inferrule{(\sigma, \sigma') \in \expr{\emph{mod}}}
  {\sigma = \sigma'}

\inferrule{(\sigma, \sigma')\in \expr{\textsf{program}(\emph{imports}, \emph{mod})}}
  {\exists \sigma''. \;(\sigma, \sigma'') \in \expr{\emph{import}},\; (\sigma'', \sigma') \in \expr{\emph{mod}}}
  
\inferrule{((\Gamma, \mathcal{C}), (\Gamma', \mathcal{C}')) \in \expr{\emph{package}\;\DOT\;\STAR}}
  {\begin{array}{c}
   \mathcal{C} = \mathcal{C}', \\ 
   I = \{u | (\emph{package}\;\DOT\;u, \tau)\in\texttt{ClassLoader}\wedge \tau <: \texttt{Module}\},\\
   I \neq \emptyset,\\
   \Gamma' = (\Gamma \setminus dom(I)) \cup I
   \end{array}
  }

\inferrule{((\Gamma, \mathcal{C}), (\Gamma', \mathcal{C}')) \in \expr{\emph{package}\;\DOT\;\emph{javaRef}}}
  {\begin{array}{c}
   \mathcal{C} = \mathcal{C}', \\ 
   (\emph{package}\;\DOT\;\emph{javaRef}, \tau) \in \texttt{ClassLoader}, \\
%   \tau = \texttt{ClassLoader}(\emph{package}\;\DOT\;\emph{javaRef}),\\
   \tau <: \texttt{Module}, \\ 
   \Gamma' = (\Gamma \setminus \emph{javaRef}) \cup \{(\emph{javaRef}, \tau)\}
   \end{array}}
\end{mathpar}\\

We introduce a utility function, \emph{isParent}, which will help us define the final rule. This function validates a possible child module in a context. 

\[\begin{array}{l}
\emph{isParent}\; : \; \textbf{St} \rightarrow\;\emph{javaRef} \rightarrow \emph{bool}\\
\emph{isParent} \;\;(\Gamma, \mathcal{C})\;\;\emph{javaRef} \;= \\
\qquad \begin{array}{rl}
 & (javaRef \in dom(\Gamma) \wedge \mathcal{C} = \bot)\\
\vee & (\emph{javaRef}, \texttt{SubModule<}\tau\texttt{>}) \in \Gamma \wedge \mathcal{C} <: \tau
\end{array}
\end{array}\]
In other words, we first check that our java reference is in our namespace. Then, we ensure that the declared parent of that java reference is an instance of the allowed super types for the module. The final rule is outlined below,\\
\begin{mathpar}
\inferrule{((\Gamma, \mathcal{C}), \sigma') \in \expr{\textsf{parent}(\emph{javaRef}, \emph{modList})}}
  {\emph{isParent}\;(\Gamma, \mathcal{C})\;\emph{javaRef}\\ \exists(\emph{javaRef}, \tau)\in\Gamma.\;\forall m\in\emph{modList}.\;(\Gamma, \tau)\in dom(|m|)}
\end{mathpar}\\
Notice that in our evaluation rules, we allow imports to shadow previously defined imports.\\

\end{section}
\end{document}